Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
a bit of cleanup

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
---

@@

results_name <- "v0.5"

@@

min_years_catch <- 25

@@

Extending this idea to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM. Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries. As a proof of concept, we first fit models to abundance indices from each stock in RAM. We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calcualted as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). There are two main reasons why an assessment model might produce poor results: model misspecification and poor data. Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(teb:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.

@@

This is not to say that the data presented here are not valuable. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained). Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet. For example, despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis. But, applying one SAR value to all stocks in a region causes innacurate estimates of stock status. The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

@@

results_name <- params$results_name

@@

min_years_catch <- params$min_years_catch

@@

Extending this idea to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM. Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries. As a proof of concept, we first fit models to abundance indices from each stock in RAM. We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calculated as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). There are two main reasons why an assessment model might produce poor results: model misspecification and poor data. Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.

@@

This is not to say that the data presented here are not valuable. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained). Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet. For example, despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis. But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status. The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
